---
title: "Hw 5"
author: "Rishab Kulkarni"
date: "12/3/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1.
```{r}
library(faraway)
data("chickwts")


w <- aov(weight ~ feed,chickwts)
summary(w)

# box-plot
plot(weight ~ feed,chickwts)

```
# 1a.
The p-value of the feed variable is 5.94e-10, making feed a significant
predictor of weights. We can say feed has an impact on weights.Thus,
there is a difference in chickens' weights based on their feed.

None of the box-plots are the same. This is more evidence that there are
differences in chickens' weights according to feed.

```{r}
# plot for model diagnostics

par(mfrow = c(2,2))
plot(w)


```

# 1b.
The residuals v. fitted values graph shows no apparent trend, so we can
claim there is homoscedasticity or constant variance in the model. The 
Normal Q-Q plot shows no major departures from normality, so we can
conclude there is normality in the model. Observation #54 is close
to the border of Cook's distance, but doesn't fall outside the dashed
line.The same goes for observation #68. No other points have this
condition, which means there are no influential points in the model.

We can double-check for homoscedasticity using Levene's test..

```{r}
# Levene's test for equality of variance

w_lev<- lm(abs(w$residuals) ~ feed,chickwts)
summary(w_lev)
```

The p-value of the F-test is 0.4324. Since this is greater than 0.01,
there is no evidence of unequal variances. This confirms our earlier
diagnosis of homoscedasticity.

# 2.
```{r}
data("infmort")


i<- aov(income ~ region,infmort)
summary(i)

# box-plot
plot(income ~ region,infmort)


```

The p-value for the region variable is 1.16e-13, making region a
significant predictor of income. We can say region has an impact on
income. Thus, income varies with region.

Again, the box-plots are varying. This is more evidence that income
varies with region.

```{r}
par(mfrow = c(2,2))
plot(i)
```

# 2a.
The residuals v. fitted values graph shows an apparent trend, so we can
claim there is no homoscedasticity or constant variance in the model. The
Normal Q-Q plot shows departures from normality, so we can claim there
is no normality in the model. The Yugoslavia observation is near the
border of Cook's distance, but isn't beyond the dashed line. No other
points have this condition, which means there are no influential points 
in the model. 

We can double-check for homoscedasticity using Levene's test..

```{r}
# Levene's test for equality of variances

i_lev <- lm(abs(i$residuals) ~ region,infmort)
summary(i_lev)
```

The p-value of the F-test is less than 0.01, which means there is evidence
of unequal variances. This confirms our earlier diagnosis of unequal
variances.

```{r}
# sample mean for each level
aggregate(income ~ region, infmort, FUN = mean)

# pairwise comparisons
pairwise.t.test(infmort$income, infmort$region, p.adj="bonferroni")


TukeyHSD(i)
```

# 2b.
In the table,the p-values less than 0.01 represent region pairs that 
are different. Thus, the pairs of regions that are different are 
Europe-Africa, Asia-Europe, and Americas-Europe.

The Tukey HSD test output shows the differences between Europe-Africa,
Asia-Europe, and Americas-Europe are significant with an adjusted p-value
of around 0.

Both the pairwise t-test and Tukey HSD test revealed identical results.

Our earlier diagnosis of the model was a lack of normality and unequal
variances, or heteroscedasticity. An appropriate remedy would be a 
transformation of the response.

# 2c.
```{r}
library(MASS)

i_bc <- boxcox(i)
lambda <- i_bc$x[which.max(i_bc$y)]
lambda

# applying transformation
i_refit <- aov(log(income) ~ region,infmort)

par(mfrow = c(2,2))
plot(i_refit)
```

We transformed the response with a box-cox power transformation. The 
optimal lambda value = -0.1414. Rounded to the nearest half, the optimal
lambda value = 0. This means a log() transformation of the response may
improve the model's fit.

After transformation of response, the Normal Q-Q plot shows increased
normality in the model. Also, there are still no influential observations
in the model. 

We can check if there's equal variance in the transformed, re-fitted model 
with the Levene's test..

```{r}
# Levene's test for equality of variances

i_refit_lev <- lm(abs(i_refit$residuals) ~ region,infmort)
summary(i_refit_lev)
```

The p-value of the F-test is greater than 0.01, which means there is
no evidence of unequal variances in the transformed, re-fitted model.
The log transformation worked; the homoscedasticity in the original model 
was remedied.

```{r}
# comparing p-values of original v. transformed models
summary(i)

summary(i_refit)
```

The p-value of region was 1.16e-13 in the original model. After the log 
transformation, the p-value of region was 7.17e-16. This means that region 
is a more significant predictor in the transformed model. Thus, the 
re-fitted model is a better fit than the original model.

```{r}
data("PlantGrowth")

p <- aov(weight ~ group,PlantGrowth)
summary(p)

# box-plot
plot(PlantGrowth$group, PlantGrowth$weight)

# Tukey HSD test
TukeyHSD(p)
```

# 3a.
The levels are ctrl, trt1, and trt2.

The box-plots are varying, implying there may be differences in weight
between the groups. In the display, the levels' box-plots exhibit varying
numerical summaries. However, further testing is required to determine if 
there are indeed differences in weight between the groups.

We can use a Tukey HSD test..

# 3b.

The Tukey HSD test output shows that only the difference between trt1 
and trt2 is significant with an adjusted p-value of 0.012. The 
differences trt1-ctrl and trt2-ctrl are non-significant with adjusted
p-values of 0.391 and 0.198, respectively.

