---
title: "HW 1"
author: "Rishab Kulkarni"
date: "9/5/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("tidyverse",repos = "http://cran.us.r-project.org")
```

## 1.

```{r}
library(tidyverse)
data("prostate", package="faraway")
summary(prostate)
```
```{r}
hist(prostate$lcavol,main='log(cancer volume)')
hist(prostate$lweight,main='log(prostate weight)')
hist(prostate$age, main='age')
hist(prostate$lbph, main='log(benign prostatic hyperplasia amount)')
hist(prostate$svi, main='seminal vesicle invasion')
hist(prostate$lcp, main='log(capsular penetration)')
hist(prostate$gleason, main= 'Gleason score')
hist(prostate$pgg45, main='percentage Gleason scores 4 or 5')
hist(prostate$lpsa, main='log(prostate specific antigen)')
```
I find it interesting that the histogram for the Gleason scores are not
normally distributed. Instead, they're skewed heavily to the right. 
It's also interesting to note that the log(capsular penetration)
histogram is heavily skewed right, whereas the log(prostate specific antigen)
is nearly a perfect normal distribution. 


![Question 2](/Users/rishabkulkarni/Downloads/Question2.pdf)

## 4.

```{r}
playbill <- read_csv('Downloads/playbill.csv')
model <- lm(CurrentWeek ~ LastWeek, data=playbill)
summary(model)
```
The fitted model is in the form of y = b_0 + b_1(x) + e.
Through our linear model summary, we have the y-intercept and the slope;
thus, we have our fitted model values. The appropriate model is
(0.9821)x + 6.805, where the response variable is the current week
and the predictor variable is the previous or last week.

## 4a.

```{r}
confint(model,level=0.95)
```
The 95% confidence interval for slope of the regression model is
(0.9514971, 1.012666). 1 is a plausible value for b_1 since it falls 
within the confidence interval.

## 4b.

```{r}
2*(1-pt(0.3217846712,16))
```
The test statistic was drawn from (est.val - hypoth. val)/(Std. error)

The pval drawn from the hypothesis test is 0.7517816,
which is larger than our alpha level 0.05; thus, we reject
the null hypothesis that b_0 = 10000.

## 4c.

```{r}
6805 + (0.9821*400000)
```
The current week box office est. from $400,000 in gross box office
in previous week is $399,645.

```{r}
predict(lm(CurrentWeek ~ LastWeek, data=playbill), newdata = data.frame(LastWeek=400000),interval= "prediction")
```
No, $450,000 is not a feasible value for the gross box office results
in the current week, since it's not within the interval (359832.8, 439442.2)

## 4d.

This rule is inappropriate; it's not reflective of our fitted regression
model which proves otherwise. There is a linear regression that models
the current week's gross box office from that of the last week's, and
it is not y = x.


![Question 5](/Users/rishabkulkarni/Downloads/Question5.pdf)
