---
title: "Hw2"
author: "Rishab Kulkarni"
date: "9/20/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("tidyverse",repos="http://cran.us.r-project.org")
install.packages("ellipse",repos="http://cran.us.r-project.org")
install.packages("ggplot2",repos="http://cran.us.r-project.org")
```

## 1a.
```{r}
library(tidyverse)
data("sat", package="faraway")
fit <- lm(total ~ expend+ratio+salary,data = sat)
summary(fit)
```
The p-val for the salary predictor is 0.0667, which is greater than our
significance level of 0.05. Thus, we fail to reject the null hypothesis that
the salary predictor is non-significant (b_salary = 0).

The p-values for the expend, ratio, and salary predictor are all above our
significance level of 0.05. Thus, we fail to reject the null hypothesis that 
the expend, ratio, and salary predictors are non-significant.

None of these predictor variables have an effect on the response, because they
all have p-values greater than 0.05; thus, they are all statistically
non-significant with no effect on the response, total.

## 1b.
```{r}
fit_takers <- lm(total~expend+ratio+salary+takers, data = sat)
summary(fit_takers)
```
The p-value for the takers predictor is 2.61e-16, which is less than our alpha
level of 0.05; thus, we can reject the null hypothesis that the takers predictor
is statistically non-significant (b_takers=0)

```{r}
anova(fit,fit_takers)
```
The linear model with the takers predictor is the preferred linear model because
according to the F-test, the corresponding p-value is 2.607e-16, which is less
than alpha level of 0.05; thus, we can reject the null hypothesis that Model 1
fits our data better than Model 2.

The F-test is equivalent to the t-test because the F-statistic is equal to 
t-test^2. Also, we get the same p-value with both tests.

## 2.
```{r}
data("prostate",package = "faraway")
fit_prostate <- lm(lpsa~lcavol+lweight+age+lbph+svi+lcp+gleason+pgg45,data= prostate)
summary(fit_prostate)
```
## 2a.
```{r}
-0.019637+ c(-1,1)*qt(0.95,88)*0.011173
-0.019637+ c(-1,1)*qt(0.975,88)*0.011173
```

## 2b.
```{r}
data("prostate",package= "faraway")
fit_significant <- lm(lpsa~lcavol+lweight+svi,data= prostate)
summary(fit_significant)
anova(fit_significant,fit_prostate)
```
Model 2 which contains all predictors is not the preferred model because its 
p-value of 0.2167 is greater than our alpha level of 0.05; thus, we fail to
reject the null hypothesis that Model 2 fits the data as well as Model 1. We 
have ample evidence that suggests Model 1 (reduced model) is the better fit
and preferred model.

## 2c.
```{r}
library(ggplot2)
library(ellipse)
CR95 = ellipse(fit_prostate, c(4,5))
CR99 = ellipse(fit_prostate, c(4,5), level=0.99)
CR998 = ellipse(fit_prostate, c(4,5), level=0.998)
myCR = rbind(CR95, CR99, CR998);
myCR = data.frame(myCR); 
names(myCR) = c("age","lbph"); 
myCR[, 'level']=as.factor(c(rep(0.95, dim(CR95)[1]), 
                              rep(0.99, dim(CR99)[1]), 
                              rep(0.998, dim(CR998)[1])));

ggplot(data=myCR, aes(x=age, y=lbph, colour=level)) + 
  geom_path(aes(linetype=level), size=1.5) + 
  geom_point(x=coef(fit_prostate)[2], y=coef(fit_prostate)[3], shape=3, size=3, colour='red') + 
  geom_point(x=0, y=0, shape=1, size=3, colour='red') 

```
The null hypothesis is that age and lbph are insignificant predictors. In other
words, their linear reg. coefficients are 0. The alternative hypothesis is that
age and lbph are significant predictors and their linear reg. coefficients
are not 0.

The location of the origin (0,0) is within the 5% level elliptical region; thus,
we fail to reject the null hypothesis that the predictors age and lbph are
insignificant to the response. We have ample evidence that suggests age and lbph
are insignificant predictors and their linear reg. coefficients are 0.

## 2d.
```{r}
n.iter = 2000; 
fstats = numeric(n.iter);
for(i in 1:n.iter){
    newprostate= prostate;
    newprostate[,c(4,3)]=prostate[sample(97),c(4,3)];
    lm = lm(lpsa ~., data= newprostate);
    fstats[i] = summary(lm)$coef[4,3]
}
length(fstats[fstats > summary(fit_prostate)$fstat[1]])/n.iter
```

## 3a.
```{r}
data("punting", package="faraway")
fit_punting <- lm(Distance~RStr+LStr+RFlex+LFlex,data=punting)
summary(fit_punting)
```
None of the predictors are statistically significant at the 95% level. All
their corresponding p-values are greater than our alpha level of 0.05.

## 3b.
```{r}
fit_punting_collective <- lm(Distance~I(RStr+LStr+RFlex+LFlex),data=punting)
anova(fit_punting_collective,fit_punting)
```
Collectively, these four predictors are significant at the alpha level 0.05.
From the F-test, the p-value of Model 2 which fits the individual predictors
is 0.5113 which is greater than 0.05. Thus, we fail to reject the null
hypothesis that Model 2 fits the data as well as Model 1 does. We have enough
evidence that suggests the four predictors collectively are significant at
the 5% alpha level.

## 3c.
```{r}
fit_punting_strength <- lm(Distance~I(RStr+LStr)+RFlex+LFlex,data=punting)
anova(fit_punting_strength,fit_punting)
```
The p-value for the full model is 0.468, which is greater than our alpha level
5%; thus, we fail to reject the null hypothesis that Model 2 (full model) is a 
better fit than Model 1 (right and left leg strength as predictor). We have
evidence that suggests Model 1 is a better fit and does not have the same effect
as our full model.

## 3d.
```{r}
library(ggplot2)
library(ellipse)
CR95 = ellipse(fit_punting, c(4,5))
CR99 = ellipse(fit_punting, c(4,5), level=0.99)
CR998 = ellipse(fit_punting, c(4,5), level=0.998)
myCR = rbind(CR95, CR99, CR998);
myCR = data.frame(myCR); 
names(myCR) = c("RStr","LStr"); 
myCR[, 'level']=as.factor(c(rep(0.95, dim(CR95)[1]), 
                              rep(0.99, dim(CR99)[1]), 
                              rep(0.998, dim(CR998)[1])));

ggplot(data=myCR, aes(x=RStr, y=LStr, colour=level)) + 
  geom_path(aes(linetype=level), size=1.5) + 
  geom_point(x=coef(fit_punting)[2], y=coef(fit_punting)[3], shape=3, size=3, colour='red') + 
  geom_point(x=0, y=0, shape=1, size=3, colour='red') 
```
The null hypothesis is that RStr and LStr are insignificant predictors. In other
words, their linear reg. coefficients are 0. The alternative hypothesis is that
RStr and LStr are significant predictors and their linear reg. coefficients
are not 0.

The location of the origin (0,0) is within the 5% level elliptical region; thus,
we fail to reject the null hypothesis that the predictors RStr and LStr are
insignificant to the response. We have evidence that suggests RStr and LStr are
insignificant predictors and their linear reg. coefficients are 0.

## 3e.
```{r}
fit_strength <- lm(Distance~RStr+LStr,data = punting)
fit_total_str <- lm(Distance~I(RStr+LStr), data=punting)
anova(fit_total_str,fit_strength)
```
The linear model that uses individual right and left leg strengths is not
sufficient to predict the response, Distance. The p-value associated with this
model is 0.5978, which is higher than our significance level 0.05. Thus, we fail
to reject the null hypothesis that Model 2 improved the fit. We have ample
evidence that suggests Model 1 (total leg strength) is a better fit.

## 5a.
```{r}
new_patient = data.frame(lcavol=1.44692,lweight=3.62301,age=65,lbph=0.3001,svi=0,lcp=-0.79851,gleason=7,pgg45=15)
predict(fit_prostate,newdata= new_patient,interval= "confidence")
```

## 5b.
```{r}
new_patient_age = data.frame(lcavol=1.44692,lweight=3.62301,age=20,lbph=0.3001,svi=0,lcp=-0.79851,gleason=7,pgg45=15)
predict(fit_prostate,newdata= new_patient_age,interval = "confidence")
```
The CI is wider because the new age value of 20 is further fromthe mean. We are 
moving away from the prostate data, so the CI will be wider.

## 5c.
```{r}
predict(fit_significant,newdata= new_patient,interval="confidence")
```
The CI is narrower; I prefer these predictions because we are using significant
predictors and thus, a better model fit. This better fit provides more accurate
predictions of our response.
